\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species

### 2024-01-24 ms.Rnw
Update of the ms file

\begin{figure}
\begin{center}
%%%%% SEAIR diagram
\resizebox{8cm}{!}{
\begin{tikzpicture}[
  square/.style={rectangle, draw=black, minimum width=0.5cm, minimum height=0.5cm, rounded corners=.1cm, fill=blue!8},
  rhombus/.style={diamond, draw=black, minimum width=0.05cm, minimum height=0.05cm, fill=purple!8,aspect = 1},
  travel/.style={circle, draw=black, minimum width=0.5cm, minimum height=0.5cm, fill=green!8},
  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=0.6cm,inner sep=0cm},
  bendy/.style={bend left=10},
  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
]
\tikzset{>={}}; % this is needed to implement the arrow redefinition

@@

\renewcommand{\arraystretch}{0.5}
\begin{table}[H]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c||}
 \hline
Parameter & Definition &Unit\\[0.5ex]
\hline\hline
$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$\\
$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$\\
$F$& Alga density& $10^5 cells \cdot L ^{-1}$\\
$P$ & Spore density & $10^3 spores \cdot L ^{-1}$\\
$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ \\
$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$ \\
$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$ \\
$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ \\
$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ \\
$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ \\
$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
$\delta$ & Sampling rate&$day ^{-1} $ \\
$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $ \\
$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ \\
$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ \\
$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ \\
$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ \\
$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ \\
$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ \\
$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ \\
$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ \\
$k_{S_n}$ & Overdispersion parameter& $Unitless$ \\
$k_{S_i}$ &Overdispersion parameter &$Unitless$ \\
$k_{I_n}$ &Overdispersion parameter &$Unitless$ \\
$k_{I_i}$ & Overdispersion parameter&$Unitless$ \\[1ex]
 \hline
\end{tabular}}
\caption{This table shows the units of the parameters of competition SIRPF2 model}
\label{Table:1}
\end{table}

@@

The adaptation of the SIRPF model to SF and SF2 as following allows for a more focused examination of the roles that resources play in ecological systems, especially when considering the competition or coexistence between different species. By isolating the food variable in the absence of parasites, these models can provide clearer insights into how food resource availability or scarcity impacts the population dynamics of native and invasive species.\\

>>>>

\begin{figure}
\begin{center}
%%%%% SEAIR diagram
\resizebox{8cm}{!}{
\begin{tikzpicture}[
  square/.style={rectangle, draw=black, minimum width=0.5cm, minimum height=0.5cm, rounded corners=.1cm, fill=blue!8},
  rhombus/.style={diamond, draw=black, minimum width=0.05cm, minimum height=0.05cm, fill=purple!8,aspect = 1,inner sep=0cm},
  travel/.style={circle, draw=black, minimum width=0.5cm, minimum height=0.5cm, fill=green!8},
  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=0.6cm,inner sep=0cm},
  bendy/.style={bend left=10},
  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
]
\tikzset{>={}}; % this is needed to implement the arrow redefinition

@@

<<Setup, include=FALSE,echo=FALSE,results=hide>>=
 load('all_shared.RData')

parms = mif.estimate
parms = as.data.frame(t(parms))

@

\renewcommand{\arraystretch}{0.5}
\begin{table}[H]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c|c||}
 \hline
Parameter & Definition &Unit & Value\\[0.5ex]
\hline\hline
$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$ & Variable\\
$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$& Variable\\
$F$& Alga density& $10^5 cells \cdot L ^{-1}$& Variable\\
$P$ & Spore density & $10^3 spores \cdot L ^{-1}$& Variable\\
$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ & \Sexpr{mysignif(parms$K_F,2)}\\
$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$rn,2)}\\
$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$ri,2)}\\
$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$  & \Sexpr{mysignif(parms$theta_In,2)}\\
$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$  & \Sexpr{mysignif(parms$theta_Ii,2)}\\
$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_Si,2)}\\
$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_Sn,2)}\\
$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_P,2)}\\
$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probn,2)}\\
$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probi,2)}\\
$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Sn,2)}\\
$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_Si,2)} \\
$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_In,2)} \\
$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Ii,2)}\\
$\delta$ & Sampling rate&$day ^{-1} $ & 0.013\\
$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $& \Sexpr{mysignif(parms$alpha,2)} \\
$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\
$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ & 0.37\\
$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ & \Sexpr{mysignif(parms$xi,2)}\\
$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$betai,2)}\\
$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$betan,2)}\\
$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSn,2)}\\
$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\
$k_{S_n}$ & Overdispersion parameter& $Unitless$ & \Sexpr{mysignif(parms$k_Sn,2)}\\
$k_{S_i}$ &Overdispersion parameter &$Unitless$ & \Sexpr{mysignif(parms$k_Si,2)}\\
$k_{I_n}$ &Overdispersion parameter &$Unitless$ & \Sexpr{mysignif(parms$k_In,2)}\\
$k_{I_i}$ & Overdispersion parameter&$Unitless$ & \Sexpr{mysignif(parms$k_Ii,2)}\\[1ex]
 \hline
\end{tabular}}
\caption{This table shows the units of the parameters of competition SIRPF2 model}
\label{Table:1}
\end{table}

@@

The adaptation of the SIRPF model to SF and SF2 as following allows for a more focused examination of the roles that resources play in ecological systems, especially when considering the competition or coexistence between different species. By isolating the food variable in the absence of parasites, these models can provide clearer insights into how food resource availability or scarcity impacts the population dynamics of native and invasive species.

### 2024-01-24 ms.tex
Update of the ms file

\renewcommand{\arraystretch}{0.5}
\begin{table}[H]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c||}
 \hline
Parameter & Definition &Unit\\[0.5ex]
\hline\hline
$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$\\
$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$\\
$F$& Alga density& $10^5 cells \cdot L ^{-1}$\\
$P$ & Spore density & $10^3 spores \cdot L ^{-1}$\\
$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ \\
$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$ \\
$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$ \\
$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ \\
$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ \\
$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ \\
$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
$\delta$ & Sampling rate&$day ^{-1} $ \\
$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $ \\
$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ \\
$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ \\
$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ \\
$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ \\
$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ \\
$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ \\
$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ \\
$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ \\
$k_{S_n}$ & Overdispersion parameter& $Unitless$ \\
$k_{S_i}$ &Overdispersion parameter &$Unitless$ \\
$k_{I_n}$ &Overdispersion parameter &$Unitless$ \\
$k_{I_i}$ & Overdispersion parameter&$Unitless$ \\[1ex]
 \hline
\end{tabular}}
\caption{This table shows the units of the parameters of competition SIRPF2 model}
\label{Table:1}
\end{table}

>>>>

\renewcommand{\arraystretch}{0.5}
\begin{table}[H]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c|c||}
 \hline
Parameter & Definition &Unit & Value\\[0.5ex]
\hline\hline
$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$ & Variable\\
$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$& Variable\\
$F$& Alga density& $10^5 cells \cdot L ^{-1}$& Variable\\
$P$ & Spore density & $10^3 spores \cdot L ^{-1}$& Variable\\
$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ & 0.092\\
$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & 0.23\\
$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & 0.23\\
$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$  & 0.66\\
$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$  & 0.37\\
$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ & 0.086\\
$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ & 0.098\\
$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ & 0.015\\
$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ & 0.92\\
$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ & 0.98\\
$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & 0.00027\\
$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& 0.00027 \\
$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& 0.0018 \\
$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & 0.000067\\
$\delta$ & Sampling rate&$day ^{-1} $ & 0.013\\
$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $& 0.011 \\
$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & 0.000030\\
$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & 0.00067\\
$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ & 0.37\\
$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ & 0.00041\\
$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ & 3.8\\
$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ & 3.2\\
$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & 0.061\\
$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ & 0.11\\
$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ & 0.45\\
$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ & 0.0018\\
$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ & 0.29\\
$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ & 0.20\\
$k_{S_n}$ & Overdispersion parameter& $Unitless$ & 4.5\\
$k_{S_i}$ &Overdispersion parameter &$Unitless$ & 5.4\\
$k_{I_n}$ &Overdispersion parameter &$Unitless$ & 1.4\\
$k_{I_i}$ & Overdispersion parameter&$Unitless$ & 1.2\\[1ex]
 \hline
\end{tabular}}
\caption{This table shows the units of the parameters of competition SIRPF2 model}
\label{Table:1}
\end{table}
